
\chapter{Layer 2 Scaling Solutions}

\section{General classification}
We classify Layer 2 scaling solution according to their setup:
\begin{itemize}
    \item \textbf{Plasma}: A tree of sidechains (chains of chains) where each user is asked to ensure the part of state it owns is available. Here, significant interaction with L1 is only required in case of dispute resolution. This came much before other solutions (proposed by Joseph Poon and Vitalik Buterin in 2017) but it was too complex to build a useful plasma.
    \item \textbf{Rollup}: An ecosystem running in parallel to a Layer 1 solution that has it's own state transition function (STF). After a bunch of transactions and blocks are created, the rollup posts transaction data / state-diff on chain and proves the combined state transition via opmistic or succinct zk proof.
    \item \textbf{Validium}: Similar to a rollup, just that the data-availability of the transaction is kept not on L1, but a separate \emph{data-availability committee} or the validium operator themselves. Validium saves on recurring cost to make data available on L1, but at the cost of security.
\end{itemize}

\section{Where sits the enourmous data?}
Whenever a layer 2 scaling solution is built, the layer 1 contract is provided with the root commitment of the updated state. But a commitment is useless unless it can be opened. A \emph{witness} links a leaf to the commitment and it requires a view of the state / transaction data for things like disputes and withdrawals.

Where does this data exists for generating a witness to do anything useful? A lot of answers can come up:

\begin{itemize}
    \item Let the data sit on the Layer 1 itself, the rollup route. This is the simplest of the solutions but seldom the most practical. Layer 1s are expensive to post data to.
    \item Let the data sit on some other special data-availability layer which has some presence on the Layer 1 concerned. For example, posting on celestia. This is the validium route. Here you still pay to post on Celestia albeit much less.
    \item Let the data sit on a data-availability layer even cheaper than celestia, maybe with a smaller quorum. This is an even cheaper validium route. But the trade-off is security, as the data-availability layer may not be as secure as for example celestia.
    \item Let the data sit with the users themselves. Let they maintain the data and post it when required. This is the plasma route. This is the most decentralized of the solutions but also the most complex to build. Also, it requires the users to be online all the time. A pretty bad UX.
\end{itemize}

\subimport{./}{plasma.tex}

\section{Bottlenecked by DA}
Layer 2 rollups have been around for a while. The first Layer 2 scaling solution was Plasma, proposed by Joseph Poon and Vitalik Buterin in 2017. The idea was to create a tree of sidechains, each of which would be able to process transactions in parallel. The root chain would be responsible for the security of the system, while the sidechains would be responsible for processing transactions.

Rollups however needs to guarantee that the state of rollup is \textbf{data-available}. Otherwise, if state is not available and the rollup-operator stops working, users cannot withdraw. As, even though the root commitment to the entire state is available, the opening are not. This makes rollups constrained by data-availability capacity of the Layer 1 chain.

Ethereum provides \texttt{375 KB} blobs every \texttt{12} seconds.

The following question hence often comes us: \emph{Can we eliminate need for state availability in L2 scalability solutions?}

\section{Rollups and levels of confirmation}

\section{Faster Layer 2 Interoperability}
In late 2020, Ethereum community started focusing on \href{https://ethereum-magicians.org/t/a-rollup-centric-ethereum-roadmap/4698}{\emph{rollup-centric roadmap}}. The idea of this was to enable various Layer 2 scaling solutions that may potentially provide a whole lot more of execution capacity than the Ethereum (henceforth to be referred as Layer1 or L1) will be capable of. 

However, Layer 2 come with following new concerns:
\begin{itemize}
    \item \textbf{Security}: The security of Layer 2's is the concern of how easy it is to compromise a L2 in comparison to an L1. Generally, in a proof-of-stake system like ethereum, a lot of security is borrowed from stake at play. As of writing, this is around 33.8 million ETH staked / around 136.8 billion USD securing the ecosystem. Can the L2 match this level of security?
    \item \textbf{Decentralization}: Control of the system especially with chains like Ethereum is fairly decentralized. As of writing, around 1 million validators run the nodes for the main chain\footnote{This number is a bit misleading given a lot is contributed by single institutions like \href{stake.fish}{stake.fish} running multiple validator nodes given each validator can only stake 32 ETH per validator. \href{https://eips.ethereum.org/EIPS/eip-7251}{The Ethereum Pectra Upgrade (EIP-7251)} is supposed to fix it with increased \texttt{MAX\_EFFECTIVE\_BALANCE} the staking balance per validator, in hopes consolidating many such large validators into fewer entities}. Layer 2 solutions hardly have decentralized sequencers and hence sit almost on the far end of decentralization argument. Ca the L2 match this level of decentralized controls? 
    \item \textbf{Fragmentation}: Every L2 is almost like a separate "room", although situated within the same "house", the L1. This eventually leads to fragmentation of both users, and the liquidity. On a DEX, liquidity is directly tied to the quantity of tokens users have committed to the liquidity pools. If a crypto asset lacks sufficient liquidity, token holders may face difficulty selling their tokens when they wish. L2 brings in another dimenstion to this problem. Not only are tokens spread thin among various DeFi apps on the same chain, they are now spread among various DeFi apps on \emph{various} chains.
\end{itemize}

The Interoperability problems is concerned with the \emph{fragmentation} issue brought in by the advent of L2s. 

In its most basic essence, Interoperability is concerned with a transaction that intends to affect state of two different L2 rollup chains. For example, a token transfer from $A$'s account on rollup $X$ to $B$'s account on rollup $Y$ is a transaction that affects the state of both $X$ and $Y$.